### 2020-03-06 sources/tools/make-tableau-data.Rmd
start on tableau test

---
title: "Make Database for Tableau Prototype"
output: html_document
---

# Overhead
```{r overhead, include = FALSE}
packages_vector <- c("tidyverse",
                     "jsonlite",
                     "visioneval")

need_to_install <- packages_vector[!(packages_vector %in% installed.packages()[,"Package"])]

if (length(need_to_install)) install.packages(need_to_install)

for (package in packages_vector) {
  library(package, character.only = TRUE)
}

# import::here(ve.export,.from="./exporter.R")
```

# Remote I/O
```{r remote-io}
viewer_dir <- "~/Documents/GitHub/VisionEval-Dev/sources/VEScenarioViewer/data/VERSPM/"
output_dir <- "./example_dir/tableau_test/"

viewer_file_name <- paste0(viewer_dir,  "verspm.js")

scenario_config_file_name <- paste0(output_dir, "example_scenario_manage.csv")

output_file_name <- paste0(output_dir, "database-for-tableau.csv")
```

# Convert Javascript data to CSV
```{r convert-java}
temp_out_name <- paste0(output_dir, "verspm.JSON")
input  <- readLines(viewer_file_name)
  step_1 <- gsub(pattern = "var data =", replace = "", x = input)
  step_2 <- str_replace(step_1, ";", "")
  writeLines(step_2, con = temp_out_name)

temp_df <- fromJSON(temp_out_name, flatten = TRUE)

write_csv(temp_df, path = paste0(output_dir, "verspm.csv"))
```

# Prepare Tableau
```{r prep-tableau}
measure_names_vector <- c("GHGReduction", 
                          "DVMTPerCapita",
                          "WalkTravelPerCapita",
                          "TruckDelay",
                          "AirPollutionEm",
                          "FuelUse",
                          "VehicleCost",
                          "VehicleCostLow")

config_df <- read_csv(scenario_config_file_name, 
                      col_types = cols(.default = col_character(), execute = col_logical()))

category_name_dict_df <- distinct(config_df, category_name, category_label) %>%
  mutate(category_label = str_replace(category_label, " ", "_")) %>%
  select(short = category_name, long = category_label)

level_name_dict_df <- distinct(config_df, category_name, level_name, level_label)

outcomes_df <- read_csv(paste0(output_dir, "verspm.csv"),
                        col_types = cols(.default = col_character())) %>%
  mutate_at(vars(measure_names_vector), as.numeric)

working_df <- outcomes_df

for (category in unique(level_name_dict_df$category_name)) {

  level_dict_df <- level_name_dict_df %>%
    filter(category_name == category) %>%
    distinct(level_name, level_label)

  new_category_name <- paste0(category, "_NAME")

  working_df <- working_df %>%
    rename(subject := !!category) %>%
    left_join(., level_dict_df, by = c("subject" = "level_name")) %>%
    mutate(subject_copy = level_label) %>%
    rename(!!new_category_name := subject_copy) %>%
    rename(!!category := subject) %>%
    select(-level_label)

}

for (category in category_name_dict_df$short) {

  long_name <- filter(category_name_dict_df, short == category)$long

  new_category_name <- paste0(category, "_NAME")

  working_df <- working_df %>%
    rename(!!long_name := !!new_category_name) 
}

output_df <- working_df %>%
  mutate(scenario_index = row_number())

```

# Write
```{r}
write_csv(output_df, path = output_file_name)
```