### 2021-05-20 sources/tools/tableau/make-tableau.Rmd
Initial files for tableau export/visualization, scenario manager
- Make output folder just 'scenarios' again; this works
- Remove land use from scenario manager file, was identical to bicycle
- Fix errors in the tutorial
- Restore seed = T in RunScenarios.R

---
title: "Make Tableau from Tara's Files"
output: html_notebook
---

# Overhead
```{r overhead, include = FALSE}
packages_vector <- c("tidyverse")

need_to_install <- packages_vector[!(packages_vector %in% installed.packages()[,"Package"])]

if (length(need_to_install)) install.packages(need_to_install)

for (package in packages_vector) {
  library(package, character.only = TRUE)
}

source("tableau_scenario_viewer.R")
```

# Remote I-O
```{r remote-io}
input_dir <- "../input/"
output_dir <- "../output/"

input_viewer_dir <- paste0(input_dir, "VERSPM/")

output_file_name <- paste0(output_dir, "for-tableau.csv")

```

# Make Scenario database
```{r}
output_df <- ve.tableau_scenario_viewer.MakeCSV(input_viewer_dir)
```

# Write Data
```{r write-data}
write_csv(output_df, path = output_file_name)
```