### 2021-09-02 sources/modules/VEModel/extensions/Model Run/Single_Scenarios_Results.Rmd
Rationalizing stage construction; starting on scenarios

---
title: "VE-RSPM Single Scenarios Results"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

<!-- Add the following line manually to the rendered HTML document so that IE does not block the javascript elements: -->
<!-- saved from url=(0014)about:internet -->

# Metropolitan Data Visualization

Visualize daily VMT for all scenarios. This is done by combining seven DVMT outputs by location and type of mode. Percent change is calculated relative to the base scenario value.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F)

# This sections sets the location of the VisionEval installation for each of the Volpe team members.
# The result is that the object `ve.runtime` is the full path to where VisionEval is installed, and the results from the single scenarios runs are located.

current_file_directory <- dirname(rstudioapi::getActiveDocumentContext()$path)
ve.runtime <- dirname((dirname(current_file_directory)))

#ve.runtime <- file.path(dirname('~/'), 'Desktop', 'VE_Tool_Test')

if(!dir.exists(ve.runtime)){
  stop('Please specify `ve.runtime` path in the setup chunk of code. \n This will be the path to your installed version of VisionEval')
}

# Make a directory to save JPEG outputs as well
fig_save = file.path(ve.runtime, 'models', 'VERSPM_Scenarios', 'Single_Scenarios_Figures')
if(!dir.exists(fig_save)) { dir.create(fig_save) }

knitr::opts_knit$set(root.dir = ve.runtime)

library(tidyverse)
library(plotly) #install.packages('plotly')
library(DT) # install.packages('DT')

overview_file = file.path(ve.runtime, 
                          'models', 'VERSPM_Scenarios',
                          'Single_Scenarios_Overview_Template.xlsx')

if(!file.exists(overview_file)){
  stop('Please copy over the file `Single_Scenarios_Overview_Template.xlsx` to your VisionEval directory, \n and place it in the `models/VERSPM_Scenarios` location')
}

# Read in scenario descriptions
scenario_desc <- readxl::read_excel(file.path(ve.runtime,
                                              'models', "VERSPM_Scenarios",
                                              'Single_Scenarios_Overview_Template.xlsx'))

# clean up trailing whitespace
scenario_values <- scenario_desc %>%
  mutate(`Model Run Name` = gsub('\\s+', '', `Model Run Name`))

# scenario_values <- scenario_values %>%
#   mutate(ModelNameLabel = ifelse(scenario_values$Level == 'NA' |
#         str_detect(scenario_values$Folder,'C0\\d'), scenario_values$`Model Run Name`,
#         paste0(scenario_values$`Model Run Name`, scenario_values$Level)))
# 
# scenario_values <- scenario_values %>%
#   mutate(`ModelNameLabel` = gsub('\\s+', '', `ModelNameLabel`))

```

## DVMT Plots

### Percent change

The following looks at percent change in each case relative to the baseline scenario for 2038.

```{r dvmt_rel}

marea_results <- read.csv(file.path(ve.runtime, 'models', 'VERSPM_Scenarios','Single_Scenarios_Metrics_Marea.csv'))

marea_results <- marea_results %>% 
  mutate(`modelName` = gsub('\\s+', '', `modelName`)) %>% 
  mutate(model_run_no = str_extract(modelName, '[A-Z]\\d{1,3}$')) %>% 
  left_join(scenario_values, by = c("modelName" = "Model Run Name"))

marea_results <- marea_results %>%   
  mutate(modelName = ifelse(modelName == 'VERSPM_base_model', 
                            '0 - Base',
                            paste(model_run_no, `Short Name`, sep = ' - ')))

#info for regional dvmt graphic

dvmt <- marea_results %>%
  select(modelName, contains('Dvmt')) %>%
  rowwise() %>%
  mutate(regionalDVMT = sum(c_across(contains('Dvmt'))))

dvmtbase = as.numeric(dvmt %>% filter(modelName == '0 - Base') %>% select(regionalDVMT))
#dvmtbase = as.numeric(dvmt %>% filter(modelName == 'VERSPM_NVTA') %>% select(regionalDVMT))

dvmt <- dvmt %>%
  mutate(regDVMTAbsDif = regionalDVMT - dvmtbase,
         regDVMTPerChange = (regDVMTAbsDif/dvmtbase)*100,
         color = ifelse(regDVMTPerChange > 0, "pos","neg"),
         pct_text = paste('Scenario:', modelName, '\n', round(regDVMTPerChange, 2), '%'),
         mi_text = paste('Scenario:', modelName, '\n', format(round(regDVMTAbsDif, 2),
                                                              scientific = F, 
                                                              big.mark = ','), 'mi/day'))

gp1 <- ggplot(dvmt, aes(x = modelName, y = regDVMTPerChange, fill = color,
                        text = pct_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Percent Change Regional DVMT", x = "Scenario", y = "Percent Change") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Save the output figure to disk for inclusion in a manuscript
ggsave(file.path(fig_save, 'DVMT_Rel_Change.jpeg'))

# And plot dynamically for viewing.
ggplotly(gp1, tooltip = 'text')
```

### Absolute change

```{r dvmt_abs}
gp2 <- ggplot(dvmt, aes(x = modelName, y = regDVMTAbsDif, fill = color,
                        text = mi_text)) +
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Absolute Change Regional DVMT (mi/day)", x = "Scenario", y = "Absolute Change")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'DVMT_Abs_Change.jpeg'))

ggplotly(gp2, tooltip = 'text')
```

## Energy Usage plots

### Truck energy usage relative change

Next we look a the relative change between scenarios for truck energy usage.

```{r truck_energy_rel}

#info for commercial heavy truck gge (sum columns k-m)
# use `matches()` to use regular expressions first to find variable names that start with
# ComSvc or HvyTrk, and then a second one to find only those which also have GGE at the end
gge <- marea_results %>%
  select(modelName, matches('^(ComSvc|HvyTrk)')) %>%
  select(modelName, matches('GGE$')) %>%
  rowwise() %>%
  mutate(ComHvyTruckGGE = sum(c_across(contains('GGE'))))

ggebase = as.numeric(gge %>% filter(modelName == '0 - Base') %>% select(ComHvyTruckGGE))
#ggebase = as.numeric(gge %>% filter(modelName == 'VERSPM_NVTA') %>% select(ComHvyTruckGGE))

gge <- gge %>%
  mutate(ComHvyTruckGGEAbsDif = ComHvyTruckGGE - ggebase,
         ComHvyTruckGGEPerChange = (ComHvyTruckGGEAbsDif/ggebase)*100,
         sign = ifelse(ComHvyTruckGGEPerChange > 0, "pos","neg"),
         pct_text = paste('Scenario:', modelName, '\n', round(ComHvyTruckGGEPerChange, 2), '%'),
         mi_text = paste('Scenario:', modelName, '\n', format(round(ComHvyTruckGGEAbsDif, 2),
                                                              scientific = F,
                                                              big.mark = ','), 'Gasoline gallone equiv.'))

# View(gge)

gp3 <- ggplot(gge, aes(x = modelName, y = ComHvyTruckGGEPerChange, fill = sign,
                       text = pct_text)) +
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Percent Change Commercial Service and  Heavy Truck GGE", x = "Scenario", y = "Percent Change")+
  theme(axis.text.x = element_text(angle = 45))

ggsave(file.path(fig_save, 'Truck_GGE_Pct_Change.jpeg'))

ggplotly(gp3, tooltip = 'text')

```

## Metropolitan area data table

```{r marea_table}
marea_table <- marea_results %>% select(-names(marea_results)[names(marea_results) %in% names(scenario_values)])

val_length <- sapply(marea_table, FUN = function(x) ifelse(is.numeric(x), max(nchar(as.character(round(x, 0)))),
                                                           max(nchar(x))))

datatable(marea_table,
          rownames = F,
          filter = 'top',
          extensions = c('Buttons','ColReorder', 'Select'),
 options = list(
   pageLength = 10,
   dom = 'BSfrtip',
   colReorder = TRUE,
   select = TRUE,
   search = list(regex = TRUE), # add regex searching to all search boxes
   buttons = c('csv', 'excel', 'colvis'),
   autoWidth = TRUE,
   scrollX = TRUE, scrollY = 600)
  ,
 caption = "Complete metropolitan area results for NVTA scenarios." ) %>%
  formatRound(which(val_length >= 5 & !grepl('model', names(val_length))), 0) %>%
  formatRound(which(val_length < 5 & !grepl('model', names(val_length))), 2)

```

# Household Data Visualization

In addition to examining the outputs which are reported at the entire metropolitan area level, we also investigated the outputs rolled up across all households.

## Energy consumption - gasoline equivalents

```{r household_prep, cache = F}
hhfile <- readr::read_csv(file.path(ve.runtime, 'models','VERSPM_Scenarios' ,'Single_Scenarios_Metrics_Hh.csv'))

# Join in the scenario names
hhfile <- hhfile %>%
  mutate(`modelName` = gsub('\\s+', '', `modelName`)) %>%
  mutate(model_run_no = str_extract(modelName, '[A-Z]\\d{1,3}$')) %>%
  left_join(scenario_values, by = c("modelName" = "Model Run Name"))

hhfile <- hhfile %>%
  mutate(modelName = ifelse(modelName == 'VERSPM_base_model',
                            '0 - Base',
                            paste(model_run_no, `Short Name`, sep = ' - ')))

# Calculate mode split
hhfile <- hhfile %>%
  rowwise() %>%
  mutate(total_trips = sum(WalkTrips, BikeTrips, TransitTrips, VehicleTrips),
         pct_Walk = WalkTrips / total_trips,
         pct_Bike = BikeTrips / total_trips,
         pct_Transit = TransitTrips / total_trips,
         pct_Vehicle = VehicleTrips / total_trips)

scenario_values <- hhfile %>%
  filter(modelName != '0 - Base') %>% # exclude the base scenario
  group_by(modelName) %>%
  summarize(sum_Hh_GGE = sum(DailyGGE),
            sum_Hh_kWh = sum(DailyKWH),
            sum_Hh_CO2e = sum(DailyCO2e),
            median_Income = median(Income),
            median_OwnCost = median(OwnCost, na.rm = T),
            median_pct_Walk = median(pct_Walk),
            median_pct_Bike = median(pct_Bike),
            median_pct_Transit = median(pct_Transit),
            median_pct_Vehicle = median(pct_Vehicle)
            )

base_values <- hhfile %>%
  filter(modelName == "0 - Base") %>%
  group_by(modelName) %>%
  summarize(base_Hh_GGE = sum(DailyGGE),
            base_Hh_kWh = sum(DailyKWH),
            base_Hh_CO2e = sum(DailyCO2e),
            median_Income = median(Income),
            median_OwnCost = median(OwnCost, na.rm = T),
            median_pct_Walk = median(pct_Walk),
            median_pct_Bike = median(pct_Bike),
            median_pct_Transit = median(pct_Transit),
            median_pct_Vehicle = median(pct_Vehicle)
            )

# We use base_values$... to compare the scenario values with the corresponding base values data frame

scenario_values <- scenario_values %>%
  mutate(GGEPerChange = ((sum_Hh_GGE - base_values$base_Hh_GGE) / base_values$base_Hh_GGE) * 100,
         KWHPerChange = ((sum_Hh_kWh - base_values$base_Hh_kWh) / base_values$base_Hh_kWh) * 100,
         CO2PerChange = ((sum_Hh_CO2e - base_values$base_Hh_CO2e) / base_values$base_Hh_CO2e) * 100,
         gge_color= ifelse(GGEPerChange > 0, "pos","neg"),
         gge_pct_text = paste('Scenario:', modelName, '\n', round(GGEPerChange, 2), '%'),
         gge_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_GGE, 2),
                                                               scientific = F,
                                                               big.mark = ','), 'Gasoline gallone equiv.'),
         kwh_color= ifelse(KWHPerChange > 0, "pos","neg"),
         kwh_pct_text = paste('Scenario:', modelName, '\n', round(KWHPerChange, 2), '%'),
         kwh_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_kWh, 2),
                                                               scientific = F,
                                                               big.mark = ','), 'kWh'),
         co2_color= ifelse(CO2PerChange > 0, "pos","neg"),
         co2_pct_text = paste('Scenario:', modelName, '\n', round(CO2PerChange, 2), '%'),
         co2_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_CO2e, 2),
                                                               scientific = F,
                                                               big.mark = ','), 'g CO[2] equivalent'))

gp4 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_GGE,
                                   text = gge_text)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional daily gasoline gallon equivalents", x = "Scenario", y = "GGE/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_GGE_Total.jpeg'))

ggplotly(gp4, tooltip = 'text')
```

```{r householde_ggepct}
gp5 <- ggplot(scenario_values, aes(x = modelName, y = GGEPerChange,
                                   text = gge_pct_text, fill = gge_color)) +
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Regional daily gasoline gallon equivalents % Change", x = "Scenario", y = "GGE/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_GGE_Percent_Change.jpeg'))

ggplotly(gp5, tooltip = 'text')
```

### Energy consumption - electricity

```{r household_kwh}
gp6 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_kWh,
                            text = kwh_text)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional KWH", x = "Scenario", y = "kWh/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_kWh_Total.jpeg'))

ggplotly(gp6, tooltip = 'text')
```

```{r household_kwh_pct}
gp7 <- ggplot(scenario_values, aes(x = modelName, y = KWHPerChange,
                            text = kwh_pct_text, fill = kwh_color)) +
  geom_bar(stat="identity") + theme_minimal() +
     scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Regional KWH Percent Change", x = "Scenario", y = "kWh/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_kWh_Percent_Change.jpeg'))

ggplotly(gp7, tooltip = 'text')
```

### Emissions

```{r household_emissions}
gp8 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_CO2e,
                            text = co2_text)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional CO[2]e", x = "Model Name", y ="g CO[2] equivalent / Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_CO2e_Total.jpeg'))

ggplotly(gp8, tooltip = 'text')
```

```{r household_emissions_pct}
gp9 <- ggplot(scenario_values, aes(x = modelName, y = CO2PerChange,
                            text = co2_pct_text, fill = co2_color)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional CO[2]e Percent Change", x = "Model Name", y ="g CO[2] equivalent / Day")+
     scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_CO2e_Percent_Change.jpeg'))

ggplotly(gp9, tooltip = 'text')

```

### Mode Choice
Mode choice across the various scenarios can also be explored.

```{r mode_choice}
# ranking by  % vehicle trips

mode_choice <- scenario_values %>%
  select(modelName, contains('median_pct')) %>%
  pivot_longer(cols = contains('median_pct'),
               names_to = 'Mode',
               values_to = 'median_pct') %>%
  mutate(Mode = gsub('median_pct_', '', Mode))

gp10 <- ggplot(mode_choice, aes(x = modelName, y = median_pct, fill = Mode)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Percent of household trips by mode", x = "Model Name", y ="Median % of trips by each mode") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Mode_Choice.jpeg'))

ggplotly(gp10)

```

### Ownership Costs

Looking at impact on household vehicle ownership costs.

```{r own_cost}
gp11 <- ggplot(scenario_values, aes(x = reorder(modelName, median_OwnCost), y = median_OwnCost)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Median vehicle ownership costs", x = "Model Name", y ="$ / household") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Ownership_costs.jpeg'))

ggplotly(gp11)
```

## Household level data table

```{r hh_table}
scenario_table <- scenario_values %>% select(names(scenario_values)[!grepl('_text', names(scenario_values)) &
                                                 !grepl('_color', names(scenario_values))])

val_length <- sapply(scenario_table, FUN = function(x) ifelse(is.numeric(x), max(nchar(as.character(round(x, 0)))),
                                                           max(nchar(x))))

datatable(scenario_table,
          rownames = F,
          filter = 'top',
          extensions = c('Buttons','ColReorder', 'Select'),
 options = list(
   pageLength = 10,
   dom = 'BSfrtip',
   colReorder = TRUE,
   select = TRUE,
   search = list(regex = TRUE), # add regex searching to all search boxes
   buttons = c('csv', 'excel', 'colvis'),
   autoWidth = TRUE,
   scrollX = TRUE, scrollY = 600)
  ,
 caption = "Summary of household level results for Single scenarios." ) %>%
  formatRound(which(val_length >= 5 & !grepl('model', names(val_length))), 0) %>%
  formatRound(which(val_length < 5 & !grepl('model', names(val_length))), 2)

```

```{r valid_table}

library(DT)

datatable(dvmt %>% select(-regDVMTAbsDif, -color, -pct_text, -mi_text) %>%
            relocate(regionalDVMT, .after = modelName),
           rownames = F,
          filter = 'top',
          extensions = c('Buttons'),
          options = c(
            pageLength = 10,
            dom = 'BSfrtip',
            colReorder = TRUE,
            select = TRUE,
            search = list(regex = TRUE), # add regex searching to all search boxes
            buttons = list('copy', 'print',
                           list(
                             extend = 'collection',
                             buttons = c('csv', 'excel', 'pdf'),
                             text = 'Download')),
            autoWidth = TRUE,
            scrollX = TRUE, scrollY = 600),
           caption = "VMT values from NVTA scenarios" ) %>%
  formatRound(2:9, 0) %>%
  formatStyle('regionalDVMT',
              backgroundColor = 'dodgerblue')

```

```{r save}
save(file = 'Final_Single_Scenarios_results.RData',
     list = c('marea_results',
              'scenario_desc',
              'scenario_values',
              'hhfile'))

# to confirm the total CO2 values from household file:
# co2compile <- hh_compiled %>% group_by(modelName) %>% summarize(sum_CO2 = sum(DailyCO2e))
# View(co2compiled)

```